Secure Composition of Robust and Optimising Compilers

### 2023-07-08 popl/main.tex
Merge sections 3 and 4

\section{Example Properties\pages{1.5}}\label{sec:compprop}

@@

\subsection{Temporal Memory Safety}

@@

\subsection{Spatial Memory Safety}

@@

\subsection{Memory Safety}

@@

\subsection{Strict Cryptographic Constant Time}

@@

\subsection{Memory Safe, Strict Cryptographic Constant Time}

@@

%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
\section{Monitors\pages{1.5}}\label{sec:monitors}
This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor--level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter--equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor--satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsection{Monitor for \glsfirst{tms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
  \typerule{tms-uninteresting}{
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
  }{tms-uninteresting}
  %
  \typerule{tms-abort}{
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
  }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

@@

\subsection{Monitor for \glsfirst{sms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{sms-uninteresting}{
% }{
%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
% }{sms-uninteresting}
% %
% \typerule{sms-abort}{
% }{
%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
% }{sms-abort}
  %
  \typerule{sms-use}{
    (\loc;\sandboxtag;m)\in\smsmonitor &
    n<m
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
  }{sms-use}
  %
  \typerule{sms-alloc}{
    (\loc;\sandboxtag;m)\notin\smsmonitor
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
  }{sms-alloc}
\end{center}
The state of the monitor for \gls{sms} is a set containing tuples of locations, control--tags, and the allocation size.
In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
The trace agreement is entirely straightforward and similar to the one fo \gls{tms}.
% \begin{center}
%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
%   %
%   \typerule{sms-alloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
%   }{sms-alloc-authentic}
%   %
%   \typerule{sms-dealloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
%   }{sms-dealloc-authentic}
%   %
%   \typerule{sms-use-authentic}{
%   }{
%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
%   }{sms-use-authentic}
%   %
%   \typerule{sms-branch-authentic}{
%   }{
%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
%   }{sms-branch-authentic}
%   %
%   \typerule{sms-binop-authentic}{
%   }{
%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
%   }{sms-binop-authentic}
%   %
%   \typerule{sms-empty-authentic}{
%   }{
%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
%   }{sms-empty-authentic}
%   %
%   \typerule{sms-abort-authentic}{
%   }{
%     \smstraceagree[]{\lightning}{\absterm}
%   }{sms-abort-authentic}
% \end{center}

@@

\subsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}

@@

\subsection{Monitor for \glsfirst{scct}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
  %
  \typerule{scct-none}{
  }{
    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
  }{scct-none}
  %
  \typerule{scct-abort}{
  }{
    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
  }{scct-abort}
\end{center}
The monitor state for the \gls{scct} monitor is completely empty, since it does not need to keep track of information.
As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
\begin{center}
  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{cct}.''}$\;$\\
  %
  \typerule{scct-low-authentic}{
  }{
    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
  }{scct-low-authentic}
  %
  \typerule{scct-high-authentic}{
  }{
    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
  }{scct-high-authentic}
  %
  \typerule{scct-empty-authentic}{
  }{
    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
  }{scct-empty-authentic}
  %
  \typerule{scct-abort-authentic}{
  }{
    \sccttraceagree[]{\lightning}{\absterm}
  }{scct-abort-authentic}
\end{center}
Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Any}$ event.

@@

\subsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}

>>>>

\section{Example Properties and their Monitors\pages{1.5}}\label{sec:compprop}

@@

\subsubsection{Temporal Memory Safety}

@@

\subsubsection{Spatial Memory Safety}

@@

\subsubsection{Memory Safety}

@@

\subsubsection{Strict Cryptographic Constant Time}

@@

\subsubsection{Memory Safe, Strict Cryptographic Constant Time}

@@

%\section{Monitors and Trace Agreements\pages{1.5}}\label{sec:monitors}
\subsection{Monitors\pages{1.5}}\label{sec:monitors}
This section introduces monitors for each of the properties presented earlier (\Cref{subsec:propdefs}).
The use of monitors is because they have some structure useful for inductive proofs, whereas traces lack this structure, thus making proofs difficult.
To lessen the burden when proving, each monitor makes use of a custom trace model that contains only the relevant information related to the property the monitor checks.
To project specification traces $\trace$ onto these monitor--level traces $\abstrace$, each property has an associated event agreement relation $\traceagree[]{\pi}{\trace}{\abstrace}$.
This relation performs the actual work of projecting the events, while the stutter--equivalence $\traceagree{\pi}{\trace}{\abstrace}$ remains the same for all properties $\pi$, yet is indexed by $\pi$ to emphasize the kind of event agreement.
With agreements, this section starts to define monitor satisfaction for traces and states that monitor satisfaction implies property satisfaction.
To this end, monitor satisfaction is defined as follows.
\bul{A specification trace $\trace$ monitor--satisfies property $\pi$} iff there exists a (final) monitor state $\monitor$ and an abstract trace $\abstrace$ such that the \iul{specification trace $\trace$ agrees with abstract trace $\abstrace$} and \oul{the initial monitor\footnote{In this paper, for all monitors, the initial monitor state is denoted as $\emptyset$.} can step to the (final) monitor state $\monitor$ with abstract trace $\abstrace$}.

@@

\subsubsection{Monitor for \glsfirst{tms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\tmsmonitor\bnfdef&\ \left\{\text{allocated}:L\times\sandboxtag,\text{freed}:L\times\sandboxtag\right\} \hspace{0.33cm}%
    \emptyset:=\ \left\{\text{allocated}:\emptyset,\text{freed}:\emptyset\right\}\\
    \mi{(Abstract\ Events)}~\absevent\bnfdef&\ \absev{\emptyevent} \mid \absev{Alloc\ }\loc\ \sandboxtag \mid \absev{Dealloc\ }\loc\ \sandboxtag \mid \absev{Use\ }\loc\ \sandboxtag \mid \absterm\\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absevent}}{,,Monitor $\tmsmonitor$ does one step to $\tmsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{tms-uninteresting}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{\emptyevent}}
% }{tms-uninteresting}
% %
% \typerule{tms-abort}{
% }{
%   \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absterm}
% }{tms-abort}
  %
  \typerule{tms-use}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor}{\absev{Use\ }\loc\ \sandboxtag}
  }{tms-use}
  %
  \typerule{tms-alloc}{
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\cup\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Alloc\ }\loc\ \sandboxtag}
  }{tms-alloc}
  %
  \typerule{tms-dealloc}{
    (\loc;\sandboxtag)\in\tmsmonitor.\text{allocated} &
    (\loc;\sandboxtag)\notin\tmsmonitor.\text{freed} \\
    \tmsmonitor[']=\left\{\text{allocated}: \tmsmonitor.\text{allocated}\setminus\left\{(\loc;\sandboxtag)\right\},\text{freed}: \tmsmonitor.\text{freed}\cup\left\{(\loc;\sandboxtag)\right\}\right\}
  }{
    \monitorcheck[]{\tmsmonitor}{\tmsmonitor[']}{\absev{Dealloc\ }\loc\ \sandboxtag}
  }{tms-dealloc}
\end{center}

@@

\subsubsection{Monitor for \glsfirst{sms}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\smsmonitor:=\ L\times\sandboxtag\times\mb{N}\hspace{0.33cm}&
    \mi{(Abstract\ Events)}~\absevent\bnfdef\ \absev{\varepsilon} \mid \absev{Alloc\ }\loc\ \sandboxtag\ n \mid \absev{Use\ }\loc\ \sandboxtag\ n \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\smsmonitor}{\smsmonitor[']}{\absevent}}{,,Monitor $\smsmonitor$ does one step to $\smsmonitor[']$ given event $\absevent$.''}$\;$\\
  %
% \typerule{sms-uninteresting}{
% }{
%   \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{\emptyevent}}
% }{sms-uninteresting}
% %
% \typerule{sms-abort}{
% }{
%   \monitorcheck[]{\smsmonitor}{\emptyset}{\absterm}
% }{sms-abort}
  %
  \typerule{sms-use}{
    (\loc;\sandboxtag;m)\in\smsmonitor &
    n<m
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor}{\absev{Use\ }\loc\ \sandboxtag\ n}
  }{sms-use}
  %
  \typerule{sms-alloc}{
    (\loc;\sandboxtag;m)\notin\smsmonitor
  }{
    \monitorcheck[]{\smsmonitor}{\smsmonitor\cup\left\{(\loc;\sandboxtag;n)\right\}}{\absev{Alloc\ }\loc\ \sandboxtag\ n}
  }{sms-alloc}
\end{center}
The state of the monitor for \gls{sms} is a set containing tuples of locations, control--tags, and the allocation size.
In comparison to the trace model of the \gls{tms} monitor, the trace model here is extended by sizing and positional information.
\Cref{tr:sms-use} performs a bounds check and \Cref{tr:sms-alloc} adds bounds information to the state of the monitor.
The trace agreement is entirely straightforward and similar to the one fo \gls{tms}.
% \begin{center}
%   \judgbox{\smstraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{sms}.''}$\;$\\
%   %
%   \typerule{sms-alloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Alloc}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Alloc\ }\loc\ n}
%   }{sms-alloc-authentic}
%   %
%   \typerule{sms-dealloc-authentic}{
%   }{
%     \smstraceagree[]{\ev{Dealloc}\ \loc;\sandboxtag;\securitytag}{\absev{Dealloc\ }\loc}
%   }{sms-dealloc-authentic}
%   %
%   \typerule{sms-use-authentic}{
%   }{
%     \smstraceagree[]{\ev{Use}\ \loc\ n;\sandboxtag;\securitytag}{\absev{Use\ }\loc\ n}
%   }{sms-use-authentic}
%   %
%   \typerule{sms-branch-authentic}{
%   }{
%     \smstraceagree[]{\ev{Branch\ }n}{\absev{\emptyevent}}
%   }{sms-branch-authentic}
%   %
%   \typerule{sms-binop-authentic}{
%   }{
%     \smstraceagree[]{\ev{Binop\ }n}{\absev{\emptyevent}}
%   }{sms-binop-authentic}
%   %
%   \typerule{sms-empty-authentic}{
%   }{
%     \smstraceagree[]{\emptyevent}{\absev{\emptyevent}}
%   }{sms-empty-authentic}
%   %
%   \typerule{sms-abort-authentic}{
%   }{
%     \smstraceagree[]{\lightning}{\absterm}
%   }{sms-abort-authentic}
% \end{center}

@@

\subsubsection{Combining \glsfirst{tms} and \glsfirst{sms} Monitors to obtain \glsfirst{ms}}\label{subsubsec:ms}

@@

\subsubsection{Monitor for \glsfirst{scct}}
\begin{gather*}
  \begin{aligned}
    \mi{(Abstract\ Store)}~\scctmonitor&:=\ \emptyset \hspace{0.5cm}
    \mi{(Abstract\ Events)}~\absevent:=\ \absev{\varepsilon} \mid \absterm \mid \absev{Any} \\
  \end{aligned}
\end{gather*}
\begin{center}
  \judgbox{\monitorcheck[]{\scctmonitor}{\scctmonitor[']}{\absevent}}{,,Monitor $\scctmonitor$ does one step to $\scctmonitor[']$ given event $\absevent$.''}$\;$\\
  %
  \typerule{scct-none}{
  }{
    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absev{\emptyevent}}
  }{scct-none}
  %
  \typerule{scct-abort}{
  }{
    \monitorcheck[]{\scctmonitor}{\scctmonitor}{\absterm}
  }{scct-abort}
\end{center}
The monitor state for the \gls{scct} monitor is completely empty, since it does not need to keep track of information.
As soon as any event is hit, the execution gets stuck, since any event is considered confidental from the perspective of this monitor.
\begin{center}
  \judgbox{\sccttraceagree[]{\event}{\absevent}}{,,Abstract event $\absevent$ is equivalent to $\event$ with respect to \gls{cct}.''}$\;$\\
  %
  \typerule{scct-low-authentic}{
  }{
    \sccttraceagree[]{\event[_b];\sandboxtag;\unlock}{\absev{\emptyevent}}
  }{scct-low-authentic}
  %
  \typerule{scct-high-authentic}{
  }{
    \sccttraceagree[]{\event[_b];\sandboxtag;\lock}{\absev{Any}}
  }{scct-high-authentic}
  %
  \typerule{scct-empty-authentic}{
  }{
    \sccttraceagree[]{\emptyevent}{\absev{\emptyevent}}
  }{scct-empty-authentic}
  %
  \typerule{scct-abort-authentic}{
  }{
    \sccttraceagree[]{\lightning}{\absterm}
  }{scct-abort-authentic}
\end{center}
Accordingly, the event agreement simply disregards all events that involved public data ($\unlock$) while mapping any other event that does involve private data ($\lock$) to the abstract $\absev{Any}$ event.

@@

\subsubsection{Combining \gls{ms} and \gls{scct} Monitors to obtain \gls{msscct}}
